---
title: "Webinar-2 October 5 2024"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```

#### Load the food texture dataset
```{r}
#file = 'http://openmv.net/file/food-texture.csv'
file = 'Data/food-texture.csv'
foodData = read.csv(?, header = ?, row.names = ?, stringsAsFactors = ?)
head(foodData, 2)
str(foodData)
```

#### Modify Crispy column to reflect high (0) or low (1) crispiness
```{r}
foodData = foodData %>% mutate(Crispylevel = ifelse(?, ?, ?))
```

##### Change Crispy and Crispylevel columns to factor (categorical) type
```{r}
# Continuous features -> Oil, Density, Hardness, Fracture
# Categorical features -> Crispy (8 levels 8 through 15) and Crispylevel (2 levels 0 and 1)
# Crispy is a categorical feature with an order
str(foodData)
categorical_cols = c(?, ?)
foodData[categorical_cols] = lapply(foodData[?], ?)
str(foodData)
```

#### Visualize the Oil Percentage feature (continuous) using a histogram
```{r}
p = ggplot(data = ?) +
  geom_?(aes(x = ?), binwidth = ?, color = ?)
p
```

#### Visualize the Oil Percentage feature (continuous) for Crispy value 10 food items using a histogram
```{r}
p = ggplot(data = ?) +
  geom_?(aes(x = ?), binwidth = ?, color = ?)
p
```

#### Scatter plot between Oil Percentage and Density (relating two continuous variables)
```{r}
p = ggplot(data = ?) +
  geom_?(aes(x = ?, y = ?), color = ?) +
  labs(x = ?, y = ?, title = ?)
p
```

#### Scatter plot between OilPercentage and Density color coded with Crispylevel
```{r}
p = ggplot(data = ?) +
  geom_?(aes(x = ?, y = ?, color = ?)) +
  labs(x = 'Density', y = 'Oil percentage', title = 'Pastry Analysis')
p
```

#### Smooth line plot using ggplot
```{r}
p = ggplot(data = ?) +
  geom_?(aes(x = ?, y = ?))
p
```

#### Plot the distribution of Crispy values (categorical variable)
```{r}
p = ggplot(data = ?) +
  geom_?(aes(x = ?))
p
```

#### Summarizing sample data: mean, variance, standard deviation, median, percentiles, and quantiles
```{r}

```

#### Visualize the Oil percentage feature using a box plot
```{r}
p = ggplot(data = ?) +
  geom_?(aes(y = ?))
p
```

#### Visualize the Oil percentage feature for different Crispylevel values using box plots
```{r}
p = ggplot(data = ?) +
  geom_?(aes(x = ?, y = ?))
p
```

#### Load the bodyfat dataset
```{r}
# Load the bodyfat dataset
file = 'Data/bodyfat.csv'
df = read.csv(file, header = TRUE, row.names = 1, stringsAsFactors = F)
str(df)
head(df, n = 2)
```

#### Create lists of categorical and continuous features
```{r}
# Create lists of categorical and continuous features
features = colnames(df)
categorical_features = c(?)
continuous_features = features[?]
```

#### Convert categorical columns to correct type
```{r}
# Convert categorical columns to correct type
df[categorical_features] = lapply(df[categorical_features], as.factor)
str(df)
```

#### Plot the distribution of males and females in the dataset
```{r}
# Plot the distribution of males and females in the dataset
p = ggplot(data = ?) +
  geom_bar(aes(x = ?), color = ?, fill = ?) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Gender',
       y = 'Frequency',
       title = 'Distribution of Genders')
p
#p+coord_flip()
```

#### Plot a component plot of weight values
```{r}
# Plot a component plot of weight values
p = ggplot(data = ?) +
  geom_point(aes(x = ?, y = ?), color = 'blue') +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot of Weights')
p
```

#### Plot a histogram of weight values
```{r}
# Plot a histogram of weight values
mean_weight = mean(df$Weight)
sd_weight = sd(df$Weight)
p = ggplot(data = df) +
  geom_histogram(aes(x = ?, y = ?),
  breaks = seq(mean_weight-3*sd_weight, mean_weight+3*sd_weight, by = 2.0),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10))+
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights')
p
```

#### Plot a histogram of height values
```{r}
```

#### Plot a histogram of bodyfat values
```{r}
```

#### Plot a histogram of weights for both genders
```{r}
# Plot a histogram of weights for both genders
p = ggplot(data = ?) +
  geom_histogram(aes(x = ?, y = ?, color = ?), alpha = ?) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights for Males and Females')
p
```

#### Plot a boxplot of bodyfat values
```{r}
# Plot a boxplot of bodyfat values
p = ggplot(data = ?) +
  geom_boxplot(aes(y = ?)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(title = 'Bodyfat Distribution')
p
```

#### Plot a boxplot of bodyfat values for both genders
```{r}
# Plot a boxplot of bodyfat values for both genders
p = ggplot(data = ?) +
  geom_boxplot(aes(x = ?, y = ?)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(title = 'Bodyfat Distribution for Males and Females')
p
```

#### Introduce a new categorical column in the dataframe called 'Weightlevel' with the following levels: (1) low (< 25th percentile of weights) (2) medium (in the IQR of weights) (3) high (>75th percentile of weights)
```{r}
# Introduce a new categorical column in the dataframe called weight level with the following levels: (1) low (< 25th percentile of weights) (2) medium (in between the 25th and 75th percentiles of weights) (3) high (> 75th percentile of weights)
df['Weightlevel'] = df['Weight'] %>% mutate
```

#### Plot a boxplot of bodyfat values for both genders across the three Weightlevel groups
```{r}
# Plot a boxplot of bodyfat values for both genders both genders across the three Weightlevel groups
````
















