---
title: "Webinar-1 September 14 2024"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries

```{r}
library(ggplot2)
library(dplyr)
```

#### Creating homeogenous datastructures: vectors and matrices (built on the fundamental datatypes character, numeric, integer, and logical)

## In R, a vector is one of the most basic data structures. It is a sequence of elements that are all of the same type. You can think of a vector as a one-dimensional array that holds data, and it plays a key role in data manipulation and analysis in R.

```{r}


num_vec <- c(1.2, 3.5, 7.8)
# print(num_vec)
num_vec2 = seq(1, 10, by = 0.5)
print(num_vec2)


int_vec <- c(1L, 2L, 3L)  # 'L' indicates the number is an integer
# print(int_vec)

int_vector = 1:10
# print(int_vector)


char_vec <- c("apple", "banana", "cherry")
# print(char_vec)

log_vec <- c(TRUE, FALSE, TRUE)
# print(log_vec)





```

```{r}



# Create a 2D matrix with 2 rows and 3 columns
matrix_2d = matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3)

# Print the matrix
print(matrix_2d)


# Create a 2D matrix filled by rows
matrix_2d_byrow <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3, byrow = TRUE)

# Print the matrix
print(matrix_2d_byrow)


matrix_2d[1, 2]  # Accesses the element in the first row, second column (returns 3)

```

```{r}

# Using rbind to create a matrix by binding rows
matrix_rbind <- rbind(c(1, 2, 3), c(4, 5, 6))
print(matrix_rbind)


# Using cbind to create a matrix by binding columns
matrix_cbind <- cbind(c(1, 4), c(2, 5), c(3, 6))
print(matrix_cbind)



```

```{r}

# Create a 3x3 matrix
mat <- matrix(1:9, nrow = 3, byrow = TRUE)

# Extract the first row as a vector
first_row_vector <- mat[1, ]

print(first_row_vector)  # Prints the vector
print(class(first_row_vector))  # Shows the data type (vector)
print(first_row_vector)


```

```{r}

# Create a sequence from 1 to 10 with a step of 0.5
num_vec2=seq(1, 10, by = 0.5)

# Create a matrix using the sequence with 4 rows and 5 columns
matrix_seq= matrix(num_vec2, nrow = 4, ncol = 5)

# Print the matrix
print(matrix_seq)

```

#### Creating a list, a heterogenous datastructure

```{r}


# Create a list with different data types
my_list <- list("R Programming", 123, TRUE, c(1, 2, 3))

# Print the list
print(my_list)
str(my_list)

# Lists are indexed by double square brackets [[ ]], and each element is displayed separately.



# Create a named list
my_list_named <- list(course = "R Programming", version = 4.0, completed = TRUE)

# Print the named list
print(my_list_named)



# Create a nested list
nested_list = list(a = 1, b = list(c = 2, d = "Nested List"))

# Access an element in the nested list
nested_list$b$c  # Returns 2




```

```{r}

# Combine two lists
list1 <- list(1, 2, 3)
list2 <- list("A", "B", "C")
combined_list <- c(list1, list2)

# Print the combined list
print(combined_list)



# Flatten a list
flattened_list <- unlist(combined_list)

# Print the flattened list
print(flattened_list)


length(combined_list)  # Returns 4
lapply(combined_list, class)  # Returns the class/type of each element in the list
sapply(combined_list, class)




```

#### Accessing elements of a datastructure

```{r}
```

#### Loading data into a dataframe, a heterogenous datastructure

```{r}



file = 'Data/food-texture.csv'
foodData = read.csv(file, header = TRUE, row.names = 1)


#str(foodData)

# Summary statistics of the dataset
# summary(foodData)

# Check for missing values
# colSums(is.na(foodData))


# Mean of 'Oil' column
mean(foodData$Oil, na.rm = TRUE)

# Median of 'Density' column
median(foodData$Density, na.rm = TRUE)





# head(foodData, 2)

```

#### Attributes of a dataframe

```{r}
```

#### Get the data type and the data structure associated with an object

```{r}
```

#### Accessing elements of a data frame in a particular column

```{r}

foodData$Oil
foodData[['Oil']] # reverse in python
foodData[, 'Oil'] # reverse in python

foodData['Oil'] # return a dataframe of foodData Dataframe.
foodData[[1]]

```

#### Accessing elements of multiple columns of a dataframe (using base R and dplyr)

```{r}

library(dplyr)

summary_stats <- foodData %>%
  summarise(
    Mean_Oil = mean(Oil, na.rm = TRUE),
    Median_Oil = median(Oil, na.rm = TRUE),
    Mean_Density = mean(Density, na.rm = TRUE),
    Median_Density = median(Density, na.rm = TRUE),
    Mean_Crispy = mean(Crispy, na.rm = TRUE),
    Mean_Fracture = mean(Fracture, na.rm = TRUE),
    Mean_Hardness = mean(Hardness, na.rm = TRUE),
    Range_Oil = range(Oil, na.rm = TRUE),
    Range_Density = range(Density, na.rm = TRUE)
  )

print(summary_stats)

```

#### For-loop and if-statement

```{r}
```

#### User-defined function

```{r}

```

#### Calling the user-defined function

```{r}
```

#### Matrix example - judges problem

```{r}
```

#### Court incorrect probability

```{r}
```

#### Accessing rows of a dataframe satisfying certain conditions

```{r}
```

#### Accessing row of a dataframe satisfying certain condition and particular columns

```{r}
```

#### Accessing rows of a dataframe satisfying certain conditions

```{r}
```

#### Modifying a column of a dataframe

```{r}
```

#### Modify Crispy column to reflect high (0) or low (1) crispiness

```{r}

```

##### Change Crispy and Crispylevel columns to factor (categorical) type

```{r}
# Continuous features -> Oil, Density, Hardness, Fracture
# Categorical features -> Crispy (8 levels 8 through 15) and Crispylevel (2 levels 0 and 1)
# Crispy is a categorical feature with an order
```
