# Variables:

## Overview

Variables in R are fundamental components used to store data that can be manipulated throughout a program. Unlike many other programming languages, R does not require explicit declaration of variables; they are created upon assignment.

## Rules for Naming Variables

When creating variable names in R, several rules must be followed:

-   **Start with a letter or a period**: Variable names must begin with a letter (a-z, A-Z) or a period (.), but if they start with a period, the next character cannot be a digit.

-   **Allowed characters**: Variable names can include letters, digits (0-9), periods (.), and underscores (\_).

-   **Case sensitivity**: Variable names are case-sensitive; for example, **`age`** and **`Age`** are considered different variables.

-   **No reserved words**: Certain keywords (e.g., **`TRUE`**, **`FALSE`**, **`NULL`**) cannot be used as variable names.

## **Examples of Valid and Invalid Variable Names**

| Valid Names  | Invalid Names |
|--------------|---------------|
| `var1`       | `1var`        |
| `my.var`     | `.2var`       |
| `data_frame` | `_data`       |
| `foo_bar`    | `TRUE`        |

## **Variable Assignment**

In **R** , variables can be assigned values using several operators:

-   Assignment Operator (`=`): Commonly used but can lead to confusion in certain contexts.

```{r}
x = 10
```

-   Leftward Assignment (\<-): Preferred method for assigning values.

```{r}
y <- "Hello"

```

-   Rightward Assignment (-\>): Less common but valid.

```{r}
"World" -> z
```

Variables can also be assigned multiple values simultaneously:

```{r}
a <- b <- c <- 10
print(a)
print(b)
print(c)
```

```{r}

10 + 20
10-20
10*20
10/20
10^20
10%%3

```

```{r}

a <- b <- c <- 10

(a+10) / b +2 * 8

```

```{r}

sqrt(16)
exp(1)
abs(-9)
ceiling(1.25)
floor(1.25)
round(1.57)

```

```{r}

max(1, 2, 3, 4)
min(1, 2, 3, 4)

```

## Numeric (Double Precision)

-   This is the default type for numbers in R.
-   It stores both integers and decimals as floating-point numbers with double precision (approximately 16 decimal digits).
-   You can assign a number using `x <- 10` or `y <- 3.14`

```{r}
## Numeric Data Type

x <- 10      # Integer, but treated as numeric (double)
y <- 3.14    # Decimal number (double)
class(x)     # Output: "numeric"
class(y)     # Output: "numeric"


```

2.  Integer

-   Integers are a specific type of numeric data, but they must be explicitly defined by adding an L after the number.
-   These are stored as 32-bit signed integers.

```{r}

z <- 10L     # Integer type
class(z)     # Output: "integer"

# Without the L, even whole numbers are treated as numeric (double).

```

3.  Floating-Point Precision

-   R uses 64-bit double-precision floating-point numbers, which allows high precision but can still result in rounding errors due to the limitations of floating-point arithmetic.

4.  Type Conversion

-   R can convert between types, such as converting a numeric value to an integer using the as.integer() function.

```{r}

a <- 10.7
b <- as.integer(a)  # Convert to integer, output will be 10

class(a)
class(b)


```

```{r}

```

